\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2019-07-17 popl-2019-journal-v/paper.tex
Move figures closer to where they're referenced.

  As discussed before, we only provide guarantees (LSE and WBCF) for components respecting certain semantic rules (reasonable use of seals).
  Components that do not satisfy these requirements are still allowed in \srccm{}, but (as we will see below) function calls in such components will not behave according to the special \srccm{} semantics that guarantees LSE and WBCF.
  To distinguish trusted components (which satisfy reasonability requirements) from adversarial components (which don't), we make use of a set of trusted addresses $\ta$.
  Any component's code memory must fall entirely within or outside of $\ta$, so every component is either trusted or adversarial.
  Some well-formedness requirements are specific for adversarial and trusted components.
  Well-formed adversarial components cannot have return seals.
  On the other hand, well-formed trusted components must make return seals available to all its calls in a way that is consistent with \stktokens{}.
  This means that each return seal is only used for one call.

  Finally, well-formed components (adversarial and trusted) can export what corresponds to closures, i.e.,  capabilities for data and code memory sealed with a closure seal.
\end{jversion}

@@

\figurename~\ref{fig:well-formed-word} defines the judgement $\vdash_{\mathrm{comp-word}}$ which specifies the words that are well-formed in a component's data memory.
The judgement is defined by three rules: \textsc{W-Data}, \textsc{W-Capability}, and \textsc{W-Sealed-Capability}.
The \textsc{W-Data} rule says all data, i.e.\ integers, are well-formed words.
The \textsc{W-Capability} rule specifies that all well-formed capabilities in data memory must at most have read and write permission.
Further, if the capability is linear, then the range of authority must be within the linear address space owned by this address.
Finally, if it is non-linear, then it must be in the non-linear address space.
The \textsc{W-Sealed-Capability} specifies that well-formed sealed capabilities in data memory are sealed with a closure seal and whatever is sealed is itself a well-formed word.
Because of the last requirement, the data memory cannot initially contain sealed code capabilities.
However, a component is free to place sealed data capabilities in memory during execution.

@@

\begin{figure}[h]
  \begin{subfigure}[h]{1.0\linewidth}
    \begin{mathpar}
  \inferrule*[Right=C-Seals]{
    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \\ [\sigma_\baddr,\sigma_\eaddr] = (\sigrets \cup \sigcloss)
  }{
    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
  }
  \and
  \inferrule*[Right=C-Instr]{
    \mscode(a) \in \ints\\
{    \begin{multlined}
      ([a \cdots a + \calllen-1] \subseteq \ta \wedge \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
      \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
    \end{multlined}
}  }{
    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
  }
  \and
  \inferrule*[Right=C-Mem]{
    \text{$\mscode$ has no hidden calls}\\
    \sigrets \mathrel{\#} \sigcloss \\
    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
    \forall a \in \dom(\mscode) \ldotp 
    \sigrets,d_\sigma(a),\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a\\
    \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
  }{
    \sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode
  }
    \end{mathpar}
    \caption{Code well-formedness.}
    \label{fig:well-formed-code}
  \end{subfigure}

  \begin{subfigure}[h]{1.0\linewidth}
    \begin{mathpar}
  \inferrule*[Right=E-Sealed-Code]{
    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} \\
    \sigma \in \sigcloss
  }{
    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
  }
  \and
  \inferrule*[Right=E-Word]{
    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} w
  }{
    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
  }
    \end{mathpar}
    \caption{Export well-formedness}
    \label{fig:well-formed-export}
  \end{subfigure}
  \caption{Well-formedness judgement for code, words, and export. $\calllen$ it the length of the call code.}
  \label{fig:well-formed-aux}
\end{figure}
\end{jversion}
\subsection{Reasonable Components}
% Reasonability
The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a component from doing something silly during execution that undermines \stktokens{}.
In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
More precisely, we expect four things of a reasonable component:
\begin{enumerate}[label=(\arabic*)]
\item It checks the stack base address before performing a call.
As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
\item It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
\item It does not leak return and closure seals or means to retrieve them.
This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
\item It does not store return and closure seals or means to get them.
By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
\end{enumerate}
\begin{jversion}
  We capture these properties in 4 definitions.
  Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly or indirectly.
  Definition~\ref{def:reasonable-pc} defines the reasonable pc's which means that if it is plugged into a configuration with a register file filled with reasonable words, then  the configuration behaves reasonably.
  Definition~\ref{def:reasonable-conf} defines the reasonable configurations which captures the four informal behavioural properties.
  Finally, definition~\ref{def:reasonable-component} lifts the notion of reasonability to components.

>>>>

  As discussed before, we only provide guarantees (LSE and WBCF) for components respecting certain semantic rules (reasonable use of seals).
  Components that do not satisfy these requirements are still allowed in \srccm{}, but (as we will see below) function calls in such components will not behave according to the special \srccm{} semantics that guarantees LSE and WBCF.
  To distinguish trusted components (which satisfy reasonability requirements) from adversarial components (which don't), we make use of a set of trusted addresses $\ta$.
  Any component's code memory must fall entirely within or outside of $\ta$, so every component is either trusted or adversarial.
  Some well-formedness requirements are specific for adversarial and trusted components.
  Well-formed adversarial components cannot have return seals.
  On the other hand, well-formed trusted components can have return seals, but must make return seals available to all its calls in a way that is consistent with \stktokens{}.
  This means that each return seal is only used for one call.

  \todo[inline]{Dominique: The following line appears to explain an arbitrary point about well-formedness: I would explain here that well-formedness imposes a quite rigid structure on well-formed components: a component's code memory may contain only data and sets of seals, while the data memory contains only data, non-sealed memory capabilities to the component's data or code memory (respecting linearity) or sealed memory or code capabilities (to the respective part of the component's memory, sealed with closure seals). A component's exports should be sealed code or data capabilities, sealed with a closure seal.
    Also: perhaps move this before the point about adversarial versus trusted components.
  }
  Finally, well-formed components (adversarial and trusted) can export what corresponds to closures, i.e.,  capabilities for data and code memory sealed with a closure seal.
\end{jversion}

@@

\begin{figure}
  \begin{subfigure}[h]{1.0\linewidth}
    \begin{mathpar}
  \inferrule*[Right=C-Seals]{
    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \\ [\sigma_\baddr,\sigma_\eaddr] = (\sigrets \cup \sigcloss)
  }{
    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
  }
  \and
  \inferrule*[Right=C-Instr]{
    \mscode(a) \in \ints\\
{    \begin{multlined}
      ([a \cdots a + \calllen-1] \subseteq \ta \wedge \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
      \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
    \end{multlined}
}  }{
    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
  }
  \and
  \inferrule*[Right=C-Mem]{
    \text{$\mscode$ has no hidden calls}\\
    \sigrets \mathrel{\#} \sigcloss \\
    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
    \forall a \in \dom(\mscode) \ldotp 
    \sigrets,d_\sigma(a),\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a\\
    \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
  }{
    \sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode
  }
    \end{mathpar}
    \caption{Code well-formedness.}
    \label{fig:well-formed-code}
  \end{subfigure}

  \begin{subfigure}[h]{1.0\linewidth}
    \begin{mathpar}
  \inferrule*[Right=E-Sealed-Code]{
    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} \\
    \sigma \in \sigcloss
  }{
    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
  }
  \and
  \inferrule*[Right=E-Word]{
    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} w
  }{
    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
  }
    \end{mathpar}
    \caption{Export well-formedness}
    \label{fig:well-formed-export}
  \end{subfigure}
  \caption{Well-formedness judgement for code, words, and export. $\calllen$ it the length of the call code.}
  \label{fig:well-formed-aux}
\end{figure}
\figurename~\ref{fig:well-formed-word} defines the judgement $\vdash_{\mathrm{comp-word}}$ which specifies the words that are well-formed in a component's data memory.
The judgement is defined by three rules: \textsc{W-Data}, \textsc{W-Capability}, and \textsc{W-Sealed-Capability}.
The \textsc{W-Data} rule says all data, i.e.\ integers, are well-formed words.
The \textsc{W-Capability} rule specifies that all well-formed capabilities in data memory must at most have read and write permission.
Further, if the capability is linear, then the range of authority must be within the linear address space owned by this address.
Finally, if it is non-linear, then it must be in the non-linear address space.
The \textsc{W-Sealed-Capability} specifies that well-formed sealed capabilities in data memory are sealed with a closure seal and whatever is sealed is itself a well-formed word.
Because of the last requirement, the data memory cannot initially contain sealed code capabilities.
However, a component is free to place sealed data capabilities in memory during execution.

@@

\end{jversion}
\subsection{Reasonable Components}
% Reasonability
The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a component from doing something silly during execution that undermines \stktokens{}.
In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
More precisely, we expect four things of a reasonable component:
\begin{enumerate}[label=(\arabic*)]
\item It checks the stack base address before performing a call.
As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
\item It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
\item It does not leak return and closure seals or means to retrieve them.
This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
\item It does not store return and closure seals or means to get them.
By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
\end{enumerate}
\begin{jversion}
  We capture these properties in 4 definitions.
  Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly or indirectly.
  Definition~\ref{def:reasonable-pc} defines the reasonable pc's which means that if it is plugged into a configuration with a register file filled with reasonable words, then  the configuration behaves reasonably.
  Definition~\ref{def:reasonable-conf} defines the reasonable configurations which captures the four informal behavioural properties.
  Finally, definition~\ref{def:reasonable-component} lifts the notion of reasonability to components.
